Geometry of Ranked Nearest Neighbour Interchange Space of Phylogenetic Trees

### 2019-12-06 rnni_convexity.tex
rm old version of
discussion of algorithms should be in the beginning

In this section we introduce three algorithms, $\findpath$ (Section~\ref{section:alg_findpath}), $\csort$ (Section~\ref{section:alg_csort}), and $\mdtree$ (Section~\ref{section:alg_mdtree}).
$\findpath$ computes a $\rnni$ path between trees.
$\csort$ computes a shortest $\rnni$ path between caterpillar trees (also known as ladder trees, see later in this section for accurate definitions).
$\mdtree$ maximises the dissimilarity to find a remote tree given an arbitrary $\rnni$ tree.
While these algorithms are interesting on their own (e.g.\ in simulation studies), they will also form an important ingredient to study the geometry of the $\rnni$ space later in this paper.

@@

% % Old version of algorithm csort that is closer to bubble sort
% \begin{algorithm}[H]
% \caption{$\csort$($T,R$)}
% \label{alg:csort}
% \begin{algorithmic}[1]
% \STATE $T':= T$, $p := [T']$
% \STATE $L_{T'}:=$ list representation of $T'$, $L_R:=$ list representation of $R$
% \FOR {$i=1,\ldots,n-2$} \label{alg:csort:line:loop}
% \STATE $j = 0$ \COMMENT{Special case: cherry}
% \IF {$L_{T'}[0]$ is last element of $L_{T'}[0],L_{T'}[1],L_{T'}[2]$ in $L_R$}
% \STATE Update $T'$: swap $L_{T'}[0]$ and $L_{T'}[2]$
% \ELSIF {$L_{T'}[1]$ is last element of $L_{T'}[0],L_{T'}[1],L_{T'}[2]$ in $L_R$}
% \STATE Update $\hat T$: swap $L_{T'}[1]$ and $L_{T'}[2]$
% \ENDIF
% \STATE $j = 2$
% \WHILE {order of $L_{T'}[j]$ and $L_{T'}[j+1]$ different in $R$}
% \STATE Update $T'$: Swap $L_{T'}[j]$ and $L_{T'}[j+1]$
% \STATE $p = p+T', j = j + 1$
% \ENDWHILE
% \ENDFOR
% \RETURN $p$
% \end{algorithmic}
% \end{algorithm}

@@

The algorithm $\mdtree$ (Maximum Distance Tree) will be an important tool for finding the radius of the $\rnni$ graph.
It computes a tree with maximum distance if the input tree is a caterpillar tree, which we will prove in Section~\ref{section:diameter}.

>>>>

In this section we introduce three algorithms, $\findpath$ (Section~\ref{section:alg_findpath}), $\csort$ (Section~\ref{section:alg_csort}), and $\mdtree$ (Section~\ref{section:alg_mdtree}).
$\findpath$ computes a $\rnni$ path between trees.
$\csort$ computes a shortest $\rnni$ path between caterpillar trees (also known as ladder trees, see later in this section for accurate definitions).
$\mdtree$ maximises the dissimilarity to find a remote tree given an arbitrary $\rnni$ tree.
While these algorithms are interesting on their own (e.g.\ in simulation studies), they will also form an important ingredient to study the geometry of the $\rnni$ space later in this paper.
For example, $\mdtree$ will be an important tool for finding the radius of the $\rnni$ graph, as the algorithms computes a tree with maximum distance if the input tree is a caterpillar tree.
We will prove this fact in Section~\ref{section:diameter}.

@@

@@
